### 2014-05-12 abstract.tex
first commit

We have developed a virtual machine for Linear Meld, a forward-chaining linear logic programming language
where programs can be executed in parallel. In Linear Meld, a program is seen as a database of logical facts and a set of derivation rules.
The database of facts is partitioned by the nodes of a graph structure which leads
to parallelism when nodes are executed simultaneously.
Due to the foundations on linear logic, rules can retract facts in a declarative and structured fashion, leading
to more expressive programs.
We present the design and implementation of the virtual machine that executes on multicores,
with a focus on thread management, code organization, fact indexing, rule execution, and database organization for efficient fact insertion,
lookup and deletion. Our results show that the virtual machine is not only
capable of scaling programs with up to 16 threads but also exhibits interesting performance results when
compared against other programming languages.