### 2014-05-12 conclusions.tex

We have presented a parallel virtual machine for executing forward-chaining linear logic programs, with particular focus
on parallel scheduling on multicores, rule execution and database organization for fast insertion,
lookup, and deletion of linear facts.
Our experimental results show that the VM is able to scale the execution of programs when run with up to 16 threads.
Although this is still a work in progress, the VM also fairs well against other programming languages.
Moreover, since LM programs are concurrent by default, we can easily get better performance from the start by executing
them with multiple threads.

In the future, we intend to take advantage of linear logic to perform whole-program optimizations,
including computing program invariants, loop detection in rules and bypass of rule priorities. For instance,
we want to detect during compile time which might be the best data structure for each predicate depending in how it
is used in the rules.