### 2014-05-12 paper.tex

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{acm_proc_article-sp}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{fancyvrb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{nth}
\usepackage[backref,pageanchor=true,plainpages=false, pdfpagelabels, bookmarks,bookmarksnumbered]{hyperref}
\input{fp-macros}

\begin{document}
\newcommand{\cmu}{\ensuremath{^\dag}}
\newcommand{\fcup}{\ensuremath{^\ddag}}
  \makeatletter
    \let\@copyrightspace\relax
      \makeatother
\numberofauthors{3} %  in this sample file, there are a *total*

\title{Design and implementation of a multithreaded virtual machine for executing linear logic programs}

\author{
\alignauthor
   Flavio Cruz \\
\affaddr{Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
\email{fmfernan@cs.cmu.edu} \\
\alignauthor Ricardo Rocha \\
\affaddr{CRACS \& INESC TEC, Faculty of Sciences, University Of Porto} \\
\affaddr{Rua do Campo Alegre, 1021/1055, 4169-007 Porto, Portugal} \\
\email{ricroc@dcc.fc.up.pt} \\
\alignauthor Seth Copen Goldstein \\
\affaddr{Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
\email{seth@cs.cmu.edu} \\
}

%\institute{Carnegie Mellon University, Pittsburgh, PA 15213, USA\\
%\email{\{fmfernan, seth\}@cs.cmu.edu}
%\and
%CRACS \& INESC TEC, Faculty of Sciences, University Of Porto\\
%Rua do Campo Alegre, 1021/1055, 4169-007 Porto, Portugal\\
%\email{\{flavioc, ricroc\}@dcc.fc.up.pt}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
\input{abstract}
\keywords{linear logic programming, virtual machine, implementation}
\end{abstract}

\section{Introduction}
\input{introduction}

\section{The LM Language}\label{lm_language}
\input{language}

\section{The Virtual Machine}\label{virtual_machine}
\input{virtual_machine}

\section{Experimental Results}\label{results}
\input{preliminary_results}

\section{Related Work}\label{related_work}
\input{related_work}

\section{Conclusions}
\input{conclusions}

\bibliographystyle{abbrv}
\bibliography{refs}
\end{document}