### 2014-07-25 conclusions.tex

\section{Conclusions}

We have presented a parallel virtual machine for executing
forward-chaining linear logic programs, with particular focus on
thread management, code organization, fact indexing, rule execution
and database organization for fast insertion, lookup, and deletion of
linear facts. Experimental results show that our VM is able to scale
the execution of programs when run with up to 16 threads.
Our results also show the importance of having an efficient indexing mechanisms
for facts. With our dynamic indexing, the VM automatically detects which predicates
need to be indexed in order to improve performance.
Due to these and other optimizations, the VM fairs relatively well against other programming languages, including
compiled languages. Moreover, since LM programs are
concurrent by default, we can easily get better performance from the
start by executing them with multiple threads.
As further work, we want to improve parallel scalability and
take advantage of linear logic to perform whole-program
optimizations, including computing program invariants, loop detection
in rules and bypass of rule priorities.

We think that our virtual machine is an excellent starting point to make
logic programming more desirable in the data-mining, machine learning and
distributed/parallel programming community. Moreover, our virtual machine can
be easily extended to execute over computer networks or to execute programs on really big datasets.
In a nutshell, LM provides a concise way to describe
graph-based algorithms that can be more easily reasoned about, a clear advantage over
competing systems.
%% For instance, we want to detect during compile time which might be the
%% best data structure for each predicate depending in how it is used in
%% the rules.