### 2014-07-25 paper.tex

\documentclass{acm_proc_article-sp}
\usepackage[utf8]{inputenc}
\usepackage{epsfig}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{subfigure}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{balance}
\usepackage{listings}
\usepackage{multicol}
\lstset{language=C, basicstyle=\ttfamily, breaklines=true, frame=lines}
\input{fp-macros}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{Design and Implementation of a Multithreaded Virtual Machine for Executing Linear Logic Programs}

\numberofauthors{3}
\author{
  \alignauthor Flavio Cruz \\
    \affaddr{Carnegie Mellon University}
    \affaddr{Pittsburgh, PA 15213, USA}\\
    \email{fmfernan@cs.cmu.edu} \\
  \alignauthor Ricardo Rocha \\
    \affaddr{CRACS \& INESC TEC}\\
    \affaddr{University of Porto}\\
    \affaddr{Rua Campo Alegre 1021/1055}\\
    \affaddr{4169-007 Porto, Portugal}\\
    \email{ricroc@dcc.fc.up.pt}
  \alignauthor Seth Copen Goldstein \\
    \affaddr{Carnegie Mellon University}
    \affaddr{Pittsburgh, PA 15213, USA}\\
    \email{seth@cs.cmu.edu} \\
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

\input{abstract}
\input{introduction}
\input{language}
\input{virtual_machine}
\input{experiments}
\input{related_work}
\input{conclusions}

\balance

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrv}
\bibliography{refs}
\end{document}