### 2014-07-25 related_work.tex

\section{Related Work}
\label{related_work}

Virtual machines are a popular technique for implementing interpreters
for high level programming languages.  Due to the increased
availability of parallel machines and distributed architectures,
several machine models have been developed with parallelism in
mind~\cite{Kara:1997:AMM:265274}.  One example of such machine is the
Parallel Virtual Machine (PVM)~\cite{Sunderam90pvm:a}, which serves as
an abstraction to program heterogeneous computers as a single
machine. Another important machine is the Threaded Abstract Machine
(TAM)~\cite{CullerGSvE93,goldstein-tr94}, which defines a
self-scheduled machine language of parallel threads where a program is
represented as conventional control flow.

Prolog, the most prominent logic programming language, has a rich
history of virtual machine research centered around the Warren
Abstract Machine (WAM)~\cite{AICPub641:1983}. 
%% The WAM offers special purpose instructions, including unification
%% instructions for different kinds of data and control flow instructions
%% to implement backtracking.  The WAM is fully sequential and uses four
%% memory areas: heap, stack, trail and the push down list.  Much
%% research has been done to improve the speed and efficiency of the
%% original WAM
%% design~\cite{Costa07demand-drivenindexing,167005,Turk-Logic:1986fk}.
Prolog is naturally parallel because several clauses for the same goal
(AND-parallelism) or all goals in a clause (OR-parallelism) can be
tried in parallel. Different abstract machines for AND-parallelism has
been developed on top of the
WAM~\cite{Hermenegildo:1986:AMB:913061,Lin:1988:AEL:900478}.  For
OR-parallelism we have several models such as: the SRI
model~\cite{Warren:1987:OEM:67683.67699}, the Argonne
model~\cite{ButlerDLOOS88}, the MUSE model~\cite{Ali:1990fk} and the
BC machine~\cite{Ali88}. While all those models have been developed
using the WAM, there are also some parallel machines totally different
from the WAM, such as the PPAM~\cite{Kacsuk:1990:EMP:533578}, which is
based on a data-flow model.