### 2014-07-25 paper.tex
Update.

\documentclass{acm_proc_article-sp}
\usepackage[utf8]{inputenc}
\usepackage{epsfig}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{subfigure}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{balance}
\usepackage{listings}
\usepackage{multicol}
\lstset{language=C, basicstyle=\ttfamily, breaklines=true, frame=lines}
\input{fp-macros}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than ACM must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from Permissions@acm.org.

PPDP '14, September 08 - 10 2014, Canterbury, United Kingdom
Copyright 2014 ACM 978-1-4503-2947-7/14/09\$15.00.
http://dx.doi.org/10.1145/2643135.2643150 }
\title{Design and Implementation of a Multithreaded Virtual Machine for Executing Linear Logic Programs}

\numberofauthors{3}
\author{
  \alignauthor Flavio Cruz \\
    \affaddr{Carnegie Mellon University}
    \affaddr{Pittsburgh, PA 15213, USA}\\
    \email{fmfernan@cs.cmu.edu} \\
  \alignauthor Ricardo Rocha \\
    \affaddr{CRACS \& INESC TEC}\\
    \affaddr{University of Porto}\\
    \affaddr{Rua Campo Alegre 1021/1055}\\
    \affaddr{4169-007 Porto, Portugal}\\
    \email{ricroc@dcc.fc.up.pt}
  \alignauthor Seth Copen Goldstein \\
    \affaddr{Carnegie Mellon University}
    \affaddr{Pittsburgh, PA 15213, USA}\\
    \email{seth@cs.cmu.edu} \\
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

\input{abstract}
\input{introduction}
\input{language}
\input{virtual_machine}
\input{experiments}
\input{related_work}
\input{conclusions}

\balance

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrv}
\bibliography{refs}
\end{document}