### 2014-08-07 paper.tex
ricroc's changes.

\documentclass{sigplanconf}
\usepackage[utf8]{inputenc}
\usepackage{epsfig}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{subfigure}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{balance}
\usepackage{listings}
\usepackage{multicol}
\lstset{language=C, basicstyle=\ttfamily, breaklines=true, frame=lines}
\input{fp-macros}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{PPDP '14}{September 08--10, 2014, Canterbury, United Kingdom}
\copyrightyear{2014} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.
\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright
%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Design and Implementation of a Multithreaded Virtual Machine for Executing Linear Logic Programs}

\authorinfo{Flavio Cruz}
           {Carnegie Mellon University\\Pittsburgh, PA 15213, USA}
           {fmfernan@cs.cmu.edu}

\authorinfo{Ricardo Rocha}
           {CRACS \& INESC TEC\\University of Porto\\Rua Campo Alegre 1021/1055\\4169-007 Porto, Portugal}
           {ricroc@dcc.fc.up.pt}

\authorinfo{Seth Copen Goldstein}
           {Carnegie Mellon University\\Pittsburgh, PA 15213, USA}
           {seth@cs.cmu.edu}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

\input{abstract}
\input{introduction}
\input{language}
\input{virtual_machine}
\input{experiments}
\input{related_work}
\input{conclusions}
\input{ack}

\balance

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.
%\bibliography{refs}

\begin{thebibliography}{35}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Ali and Karlsson(1990)]{Ali:1990fk}
K.~Ali and R.~Karlsson.
\newblock Full prolog and scheduling or-parallelism in muse.
\newblock \emph{International Journal of Parallel Programming}, 19\penalty0
  (6):\penalty0 445--475, 1990.
\newblock ISSN 0885-7458.
\newblock \doi{10.1007/BF01397627}.
\newblock URL \url{http://dx.doi.org/10.1007/BF01397627}.

\bibitem[Ali(1988)]{Ali88}
K.~A.~M. Ali.
\newblock Or-parallel execution of prolog on bc-machine.
\newblock In R.~A. Kowalski and K.~A. Bowen, editors, \emph{ICLP/SLP}, pages
  1531--1545. MIT Press, 1988.
\newblock ISBN 0-262-61056-6.
\newblock URL \url{http://dblp.uni-trier.de/db/conf/iclp/iclp88.html#Ali88}.

\bibitem[Ashley-Rollman et~al.(2007)Ashley-Rollman, Rosa, Srinivasa, Pillai,
  Goldstein, and Campbell]{ashley-rollman-derosa-iros07wksp}
M.~P. Ashley-Rollman, M.~D. Rosa, S.~S. Srinivasa, P.~Pillai, S.~C. Goldstein,
  and J.~D. Campbell.
\newblock Declarative programming for modular robots.
\newblock In \emph{Workshop on Self-Reconfigurable Robots/Systems and
  Applications at IROS 2007}, 2007.

\bibitem[Ashley-Rollman et~al.(2009)Ashley-Rollman, Lee, Goldstein, Pillai, and
  Campbell]{ashley-rollman-iclp09}
M.~P. Ashley-Rollman, P.~Lee, S.~C. Goldstein, P.~Pillai, and J.~D. Campbell.
\newblock A language for large ensembles of independently executing nodes.
\newblock In \emph{International Conference on Logic Programming (ICLP)}, 2009.

\bibitem[Baelde(2012)]{Baelde:2012:LGF:2071368.2071370}
D.~Baelde.
\newblock Least and greatest fixed points in linear logic.
\newblock \emph{ACM Transactions on Computational Logic}, 13\penalty0
  (1):\penalty0 1--44, 2012.

\bibitem[Betz and Fr{\"u}hwirth(2005)]{Betz:2005kx}
H.~Betz and T.~Fr{\"u}hwirth.
\newblock A linear-logic semantics for constraint handling rules.
\newblock In \emph{Principles and Practice of Constraint Programming - CP
  2005}, volume 3709 of \emph{Lecture Notes in Computer Science}, pages
  137--151. 2005.

\bibitem[Butler et~al.(1988)Butler, Disz, Lusk, Olson, Overbeek, and
  Stevens]{ButlerDLOOS88}
R.~Butler, T.~Disz, E.~L. Lusk, R.~Olson, R.~A. Overbeek, and R.~L. Stevens.
\newblock Scheduling or-parallelism: An argonne perspective.
\newblock In R.~A. Kowalski and K.~A. Bowen, editors, \emph{ICLP/SLP}, pages
  1590--1605. MIT Press, 1988.
\newblock ISBN 0-262-61056-6.
\newblock URL
  \url{http://dblp.uni-trier.de/db/conf/iclp/iclp88.html#ButlerDLOOS88}.

\bibitem[Colmerauer and Roussel(1993)]{Colmerauer:1993:BP:154766.155362}
A.~Colmerauer and P.~Roussel.
\newblock The birth of prolog.
\newblock In \emph{The Second ACM SIGPLAN Conference on History of Programming
  Languages}, pages 37--52, New York, NY, USA, 1993.

\bibitem[Costa et~al.(2011)Costa, Damas, and
  Rocha]{DBLP:journals/corr/abs-1102-3896}
V.~S. Costa, L.~Damas, and R.~Rocha.
\newblock The yap prolog system.
\newblock \emph{CoRR}, abs/1102.3896, 2011.

\bibitem[Cruz et~al.(2014)Cruz, Rocha, Goldstein, and Pfenning]{cruz-iclp14}
F.~Cruz, R.~Rocha, S.~Goldstein, and F.~Pfenning.
\newblock {A Linear Logic Programming Language for Concurrent Programming over
  Graph Structures}.
\newblock \emph{Journal of Theory and Practice of Logic Programming, 30th
  International Conference on Logic Programming (ICLP 2014), Special Issue},
  14\penalty0 (4 \& 5):\penalty0 493--507, July 2014.

\bibitem[Culler et~al.(1993)Culler, Goldstein, Schauser, and von
  Eicken]{CullerGSvE93}
D.~E. Culler, S.~C. Goldstein, K.~E. Schauser, and T.~von Eicken.
\newblock {TAM --- a compiler controlled threaded abstract machine}.
\newblock \emph{Journal of Parallel and Distributed Computing}, 18:\penalty0
  347--370, July 1993.
\newblock URL \url{http://www.cs.cmu.edu/~seth/papers/CullerGSvE93.pdf}.

\bibitem[Girard(1987)]{girard-87}
J.-Y. Girard.
\newblock Linear logic.
\newblock \emph{Theoretical Computer Science}, 50\penalty0 (1):\penalty0
  1--102, 1987.

\bibitem[Goldstein(1994)]{goldstein-tr94}
S.~C. Goldstein.
\newblock The implementation of a threaded abstract machine.
\newblock Technical Report UCB/CSD-94-818, EECS Department, University of
  California, Berkeley, 1994.
\newblock URL \url{http://www.cs.cmu.edu/~seth/papers/goldstein-tr94.pdf}.

\bibitem[Gonzalez et~al.(2009)Gonzalez, Low, and
  Guestrin]{Gonzalez+al:aistats09paraml}
J.~Gonzalez, Y.~Low, and C.~Guestrin.
\newblock Residual splash for optimally parallelizing belief propagation.
\newblock In \emph{Artificial Intelligence and Statistics (AISTATS)}, 2009.

\bibitem[Gupta et~al.(2001)Gupta, Pontelli, Ali, Carlsson, and
  Hermenegildo]{Gupta:2001:PEP:504083.504085}
G.~Gupta, E.~Pontelli, K.~A.~M. Ali, M.~Carlsson, and M.~V. Hermenegildo.
\newblock Parallel execution of prolog programs: {A} survey.
\newblock \emph{ACM Transactions on Programming Languages and Systems
  (TOPLAS)}, 23\penalty0 (4):\penalty0 472--602, 2001.

\bibitem[Hermenegildo(1986)]{Hermenegildo:1986:AMB:913061}
M.~V. Hermenegildo.
\newblock \emph{An Abstract Machine Based Execution Model for Computer
  Architecture Design and Efficient Implementation of Logic Programs in
  Parallel}.
\newblock PhD thesis, 1986.
\newblock AAI8700203.

\bibitem[Hodas and Miller(1994)]{Hodas94logicprogramming}
J.~S. Hodas and D.~Miller.
\newblock Logic programming in a fragment of intuitionistic linear logic.
\newblock \emph{Information and Computation}, 110:\penalty0 32--42, 1994.

\bibitem[Holzbaur et~al.(2004)Holzbaur, de~la Banda, Stuckey, and
  Duck]{DBLP:journals/corr/cs-PL-0408025}
C.~Holzbaur, M.~J.~G. de~la Banda, P.~J. Stuckey, and G.~J. Duck.
\newblock Optimizing compilation of constraint handling rules in hal.
\newblock \emph{CoRR}, cs.PL/0408025, 2004.

\bibitem[Isard et~al.(2007)Isard, Budiu, Yu, Birrell, and
  Fetterly]{Isard:2007:DDD:1272996.1273005}
M.~Isard, M.~Budiu, Y.~Yu, A.~Birrell, and D.~Fetterly.
\newblock Dryad: distributed data-parallel programs from sequential building
  blocks.
\newblock In \emph{European Conference on Computer Systems (EuroSys)}, pages
  59--72, 2007.

\bibitem[Kacsuk(1990)]{Kacsuk:1990:EMP:533578}
P.~Kacsuk.
\newblock \emph{Execution Models of PROLOG for Parallel Computers}.
\newblock MIT Press, Cambridge, MA, USA, 1990.
\newblock ISBN 0262111497.

\bibitem[Kara et~al.(1997)Kara, Davy, Goodeve, and Nash]{Kara:1997:AMM:265274}
M.~Kara, J.~R. Davy, D.~Goodeve, and J.~Nash, editors.
\newblock \emph{Abstract Machine Models for Parallel and Distributed
  Computing}.
\newblock IOS Press, Amsterdam, The Netherlands, The Netherlands, 1997.
\newblock ISBN 90-5199-267-X.

\bibitem[Lam and Sulzmann(2010)]{DBLP:journals/corr/abs-1006-3039}
E.~S.~L. Lam and M.~Sulzmann.
\newblock Concurrent goal-based execution of constraint handling rules.
\newblock \emph{CoRR}, abs/1006.3039, 2010.

\bibitem[Lin and Kumar(1988)]{Lin:1988:AEL:900478}
Y.~Lin and V.~Kumar.
\newblock And-parallel execution of logic programs on a shared memory
  multiprocessor:a summary of results.
\newblock Technical report, Austin, TX, USA, 1988.

\bibitem[Liu(1998)]{Liu98extendingdatalog}
M.~Liu.
\newblock Extending datalog with declarative updates.
\newblock In \emph{International Conference on Database and Expert System
  Applications (DEXA)}, volume 1873, pages 752--763, 1998.

\bibitem[Loo et~al.(2006)Loo, Condie, Garofalakis, Gay, and
  Hellerstein]{Loo-condie-garofalakis-p2}
B.~T. Loo, T.~Condie, M.~Garofalakis, D.~E. Gay, and J.~M. Hellerstein.
\newblock Declarative networking: Language, execution and optimization.
\newblock In \emph{International Conference on Management of Data (SIGMOD)},
  pages 97--108, 2006.

\bibitem[L\'{o}pez et~al.(2005)L\'{o}pez, Pfenning, Polakow, and
  Watkins]{Lopez:2005:MCL:1069774.1069778}
P.~L\'{o}pez, F.~Pfenning, J.~Polakow, and K.~Watkins.
\newblock Monadic concurrent linear logic programming.
\newblock In \emph{Proceedings of the 7th ACM SIGPLAN International Conference
  on Principles and Practice of Declarative Programming}, PPDP '05, pages
  35--46, New York, NY, USA, 2005.

\bibitem[Low et~al.(2010)Low, Gonzalez, Kyrola, Bickson, Guestrin, and
  Hellerstein]{GraphLab2010}
Y.~Low, J.~Gonzalez, A.~Kyrola, D.~Bickson, C.~Guestrin, and J.~M. Hellerstein.
\newblock Graphlab: {A} new framework for parallel machine learning.
\newblock In \emph{Conference on Uncertainty in Artificial Intelligence (UAI)},
  pages 340--349, 2010.

\bibitem[Malewicz et~al.(2010)Malewicz, Austern, Bik, Dehnert, Horn, Leiser,
  and Czajkowski]{Malewicz:2010:PSL:1807167.1807184}
G.~Malewicz, M.~H. Austern, A.~J.~C. Bik, J.~C. Dehnert, I.~Horn, N.~Leiser,
  and G.~Czajkowski.
\newblock Pregel: a system for large-scale graph processing.
\newblock In \emph{International Conference on Management of Data (SIGMOD)},
  pages 135--146, 2010.

\bibitem[Miller(1985)]{Miller85anoverview}
D.~Miller.
\newblock An overview of linear logic programming.
\newblock In \emph{in Computational Logic}, pages 1--5, 1985.

\bibitem[Ramakrishnan and Ullman(1993)]{Ramakrishnan93asurvey}
R.~Ramakrishnan and J.~D. Ullman.
\newblock A survey of research on deductive database systems.
\newblock \emph{Journal of Logic Programming}, 23:\penalty0 125--149, 1993.

\bibitem[Sunderam(1990)]{Sunderam90pvm:a}
V.~S. Sunderam.
\newblock Pvm: A framework for parallel distributed computing.
\newblock \emph{Concurrency: Practice and Experience}, 2:\penalty0 315--339,
  1990.

\bibitem[{van Rossum}(1995)]{vanRossum:1995:PRM}
G.~{van Rossum}.
\newblock {Python} reference manual.
\newblock Report CS-R9525, Centrum voor Wiskunde en Informatica, Amsterdam, the
  Netherlands, Apr. 1995.
\newblock URL \url{http://www.python.org/doc/ref/ref-1.html}.

\bibitem[Warren(1983)]{AICPub641:1983}
D.~H.~D. Warren.
\newblock An abstract prolog instruction set.
\newblock Technical Report 309, AI Center, SRI International, 333 Ravenswood
  Ave., Menlo Park, CA 94025, Oct 1983.

\bibitem[Warren(1987)]{Warren:1987:OEM:67683.67699}
D.~H.~D. Warren.
\newblock Or-parallel execution models of prolog.
\newblock In \emph{II and Colloquium on Functional and Logic Programming and
  Specifications (CFLP) on TAPSOFT '87: Advanced Seminar on Foundations of
  Innovative Software Development}, pages 243--259, New York, NY, USA, 1987.
  Springer-Verlag New York, Inc.
\newblock ISBN 0-387-17611-X.
\newblock URL \url{http://dl.acm.org/citation.cfm?id=67683.67699}.

\bibitem[Watkins et~al.(2004)Watkins, Cervesato, Pfenning, and
  Walker]{Watkins:2004uq}
K.~Watkins, I.~Cervesato, F.~Pfenning, and D.~Walker.
\newblock A concurrent logical framework: The propositional fragment.
\newblock In \emph{Types for Proofs and Programs}, volume 3085 of \emph{Lecture
  Notes in Computer Science}, pages 355--377. 2004.
\newblock \doi{10.1007/978-3-540-24849-1_23}.

\end{thebibliography}

\end{document}